### 2016-05-13 paper.tex
initial commit

\documentclass[a4paper, 11pt, titlepage]{report}

\usepackage[margin=2.0cm]{geometry}

\usepackage{xspace}
\def \iZc {I\textsuperscript{2}C\xspace}
\def \vdd {V\textsubscript{DD}}

% Indent the first line of the first paragraph like the first lines of subsequent paragraphs
\usepackage{indentfirst}

% Don't indent any first lines
\setlength{\parindent}{0mm}

% A little space between paragraphs
\setlength{\parskip}{1.5mm}

% Display
% 1. <chapter title>
% instead of
% Chapter 1
% <chapter title>
\makeatletter
\renewcommand{\@makechapterhead}[1]{%
\vspace*{50 pt}%
{\setlength{\parindent}{0pt} \raggedright \normalfont
\bfseries\Huge\thechapter.\ #1
\par\nobreak\vspace{40 pt}}}
\makeatother

% In order to include images
\usepackage{graphicx}

% Label figures 1, 2, 3, 4... Instead of 1.1, 1.2, 2.1, 3.1 etc.
\usepackage{chngcntr}
\counterwithout{figure}{chapter}

\usepackage{siunitx}

% To have text wrapped around images 
\usepackage{wrapfig}

% Some parameters for wrapfig
\setlength{\intextsep}{0pt}%
\setlength{\columnsep}{15pt}%

% Same as regular math but with a bit of space before and after
\newenvironment{ownmath}
{\vspace{2mm}\hspace{15pt}\begin{math}}
{\end{math}\vspace{2mm}}

\begin{document}

\title{Development of a 3D display}
\author{Balduin Dettling}

% Generate title
\maketitle

% Introduction
% * after chapters to not number them
% TODO: don't show page numbers
\chapter*{Preface}

\pagenumbering{Roman}

\section*{Motivation}

For as long as I can remember, I've been fascinated by all kinds of technology. 
It was clear to me that my final thesis would be a practical one in the field of engineering.
However, I didn't have anything specific in mind for quite some time.

One day, I stumbled upon a project report on the internet about a so-called persistence of 
vision display. It was a simple row of LEDs, driven by a microcontroller, and mounted on an 
electric motor. The contraption was (and presumably still is) able to draw images into the 
air by precisely controlling the row of LEDs. Fascinated by this, I did some research and 
found that a plethora of similar projects exists, some even with detailed instructions on how 
to build the device in question.

I had the idea of building several rows of LEDs and arranging them on top of each other. 
That way, I could display several two-dimensional bitmap images above another, resulting in a
three-dimensional image.

Having had no previsous experience in the field of electronics, it seemed a bit daunting to 
build a 3D display. I could, however, motivate myself to attempt it anyway by telling myself
that the project wouldn't be much more complex than all of the 2D versions out there (which 
turned out to be not very far from the truth).

\section*{Credits}

First and foremost, I would like to express my gratitude to Mr. Patrick Spengler, my physics
teacher and the mentor of this thesis. Not only did he solve numerous malfunctions, but he also
is partly responsible for my interest in electronics and engineering.

I am thankful to Ms. Ursula Schamberger, teacher of the woodworking class, where I built the
wooden frame. Likewise, I would like to thank Mr. Hanspeter Rieder, who kindly granted me
access to the school's electronics workshop.

Last, but most certainly not least, I am indebted to Prof. Heinz Domeisen. He mentored the
advancement and refinement of the thesis in preparation for the national competition organised
by "Swiss Youth in Science".

\section*{Declaration of authorship}

I hereby declare that the following thesis has been authored entirely by myself, Balduin
Dettling, and confirm that all external sources of information are cited correctly.

\vspace{10mm}
\noindent Location and Date: \hspace{0.35\textwidth} Signed:\\
\rule[3mm]{\textwidth}{0.5pt}

% Generate table of contents
\tableofcontents

\chapter{Introduction}

\pagenumbering{arabic}

\section{Current State of The Art}

% todo: references
As discussed before, there is no shortage of projects that are similar to mine, but only
display 2D images. The name "POV display", short for persistence of vision display, has become
commonplace. Another popular name is "Propeller Clock" --- Propeller because it rotates like one,
and clock because the round shape is ideal for displaying an analog clock.

As far as I know, however, only two people have built a working version of a three-dimensional
POV display so far (and have published it on the internet).\footnote{It turned out that there was
a third one: When I posted a picture of my project on the internet, a reddit user told me about
the display he built years before.}
Neither of these is being produced on a grand scale, instead they are --- like most 2D versions
--- side projects of makers or engineers.

\section{Goals}

The objective of this project is to build a working prototype of a 3D persistence of vision
display and to document the process of doing so. The necessary steps are (roughly) as follows:

\begin{itemize}
	\item Understand and explain how and why such a display can work in principle
	\item Come up with realistic technical specifications, founded on the previously outlined 
	theory  
	\item Select parts and build the device according to the specifications
	\item Write a program that makes the device display a 3D image
	\item Draw conclusions: Did I achieve my goals? What could have been done better? 
	How can the project be further improved?
\end{itemize}

\chapter{Theory and Planning}

\section{Principle of Operation}

\subsection{Persistence of Vision}

Persistence of Vision refers to the optical illusion that results whenever the image we look
at quickly changes. In such situations, we don't immediately see the new image, instead we
continue seeing the old image for a short timespan after it has vanished. If wikipedia is to 
be believed, this timespan is about $\frac{1}{25}$ of a second, but of course it is going to 
be different for each pair of eyes.

Although we don't normally see this effect in action, there are situations where it becomes
apparent. For example, if a car drives by at nighttime, we see traces of light behind the
actual car lights for a short moment (Assuming our eyes don't follow the car and stay still
relative to the environment). In general, bright flashes of light followed by (relative)
darkness tend to be the most pronounced manifestation of persistence of vision. It is exactly 
this situation that can be recreated and exploited for the successful operation of a persistence
of vision display.

\subsection{Exploiting Persistence of Vision}

Normally, Persistence of Vision is nothing more than an interesting effect. However, we can 
use this weakness of the human eye to our benefit.

If a light source is spun quickly enough in a circle, we don't see a spinning
point anymore, but instead we see one solid circle. This is because the rotational period is
shorter than the time it takes for the light to fade away on our retinas. The rotational period
at which we stop seeing a flickering circle and start seeing a solid, still circle is called
the flicker fusion threshold.

By controlling this light source quickly and accurately, we can make it so it is always turned 
on at certain points in the circle, and always turned off at others. By doing so, we are
effectively multiplexing a zero-dimensional display (a point-like light source, e.g. an LED)
so that it displays a one-dimensional image. By rotating a whole row of LEDs (a one-dimensional
display), we can display two-dimensional images. Finally, if there are multiple rows of LEDs
that together form a two-dimensional display, it is possible to display a three-dimensional
image.

\section{Specifications}

\begin{itemize}

\item Resolution of 16 * 100 * 10 pixels (radius * circumference * height)
\item Each pixel consists of an RGB LED
\item Colour depth of 3 bit per pixel, or one bit per LED
\item Rotational frequency of at least 30 Hz, or as fast as it needs to be in order to not flicker

\end{itemize}

\section{Components}

\subsection{Control}

For controlling the LEDs, I have bought a Teensy 3.1 microcontroller development board. It has a
small footprint --- ideal for the fast rotation --- and still packs quite some power: The processor
runs at 96 MHz at a register width of 32 bit. There are 34 configurable I/O pins.

The controller has to have some kind of reference point, so that it can calculate the current speed
and angular position of the device. For this, I used a simple hall sensor on the rotating side, 
which passes by a stationary magnet after each rotation.

\subsection{Data Transmission}

34 I/O pins are quite a lot, but not quite enough to directly control 480 LEDs. In order to still
access all those LEDs individually, I had to use LED drivers. These come in the form of integrated
circuits and receive some amount of data that they use to control a corresponding amount of LEDs.

Initially, I had planned to implement a colour depth of 8 bits, allowing for 
$2^{24} = 16777216$ different colours. However, there were several difficulties that I would have
had to deal with. First of all, all of the available LED drivers had one or both of the following
shortcomings:

\begin{itemize}

\item Controlled using \iZc  instead of SPI

Because \iZc uses pullup resistors instead of a push-pull drive, it takes a little moment for the
logic level to return to its normal state. In order to ensure correct data transmission, it is 
necessary to set a relatively slow clock speed. The two \iZc interfaces of the Teensy only support
clock speeds of 2.4 MHz --- and because \iZc requires some overhead data to send out the slave's
addresses, the actual data rate will be even lower.

However, the minimum data rate required to control the LEDs at an 8-bit colour depth is as follows:

\begin{ownmath}
% Yes it's a mess but \si makes it a unit, so it is not in cursive while still being small enough
24\,\frac{\si{bit}}{\si{pixel}} * 16000\,\frac{\si{pixels}}{\si{rotation}} * 30\,\frac{\si{rotations}}{\si{second}} = \SI{11.52}{Mbps} 
\end{ownmath}

This means that I'd need at least 5 parallel \iZc lines to control all the LEDs with 8 bit colour
depth. Because each Teensy only has two of them, I'd need three Teensy boards, which would be
possible but very cumbersome.

\item Insufficient PWM frequency

If the colour depth is more than one bit, the LEDs will have to be dimmed using PWM. Normally, a
rather low frequency of a few hundred to a few thousand Hz is sufficient as a PWM frequency, and
most LED drivers modulate at frequencies within that range. In my case, however, the LEDs are 
moving into a new pixel 3000 times a second (assuming rotationat 30 Hz). This means that a PWM
frequency under 3000 Hz will involve losing information, and one slightly above 3000 Hz still
won't look good. There would have to be a significant number of PWM cycles within each pixel,
for example 10, which would require a PWM frequency of 30 KHz.

\end{itemize}

Because of that, I ditched the idea of 8-bit colour depth and instead went for the simpler approach
of one-bit colour depth. Both problems are solved by switching to 1 bit: The data rate doesn't have
to be as high (and there are lots of chips that are controlled by SPI), and PWM is not necessary
anymore.

% todo: Find out if this is necessary in the end
\newpage
\subsection{LED drivers}

\begin{wrapfigure}{r}{0cm}

\includegraphics[scale=0.35]{./images/tlc-pins.png}
\caption{Pin layout of the TLC5927 LED driver}

\end{wrapfigure}

After quite some searching around, I decided to use the TLC5927 by Texas Instruments. This is
essentially a shift register with some extra features. Like every shift register, it has a serial
data input (SDI), a serial data output (SDO), and a clock pin (CLK). Additionally, there is a latch
pin (LE), which serves the purpose of only refreshing the state of the LEDs once all data has been
sent.

Furthermore, there is a pin called "output enable" (OE). If it is at GND, all LEDs are turned off,
and it's at \vdd, they are controlled by the data that the driver chip has received.

The most important feature, however, is the fact that each output is controlled by a constant
current driver. Without these, it would've been necessary to solder a resistor in series
with each individual LED, for a total of 480 resistors, 960 additional solder joints and a whole
lot of wasted board space. With the constant current drivers, however, I could use a single
resistor to set the current for all 16 LEDs attached to one chip.

Unlike most simple shift registers, the drivers' outputs are current sinks instead of current 
sources. This will be important when selecting the LEDs: In order to control each colour separately 
in an RGB LED, the cathodes have to be separated, while the anodes can be connected together.

\subsection{LEDs}

Like all the other parts, I ordered the LEDs from Digikey. Their website
contains an excellent tool, enabling the user to restrict the selection based on different criteria.
My requirements were the following:

\begin{itemize}

\item Colours: Red, green and blue
\item Current: 20 mA or more. Because the LEDs are within one pixel only $\frac{1}{100}$ of the
time, they are essentially pulse width modulated, reducing their effective brightness. In order to
still see the image clearly without darkening the room, they have to be bright enough.
\item Mounting type: Surface mount, so that I can route traces past the LEDs on the other side 
of the PCB
\item Size: At most 5 * 5 mm, preferably smaller, but still large enough to be able to solder them
by hand
\item Diffused lens, so the colours mix well even when viewing the LED directly instead of using
it for illumination.

\end{itemize}

These requirements reduced the category "LED Indication - Discrete" from 18940 to 130 products.
From those remaining products, I chose the one that was cheapest in a quantity of 160. This 
led me to an LED with the beautiful name CLVBA-FKA-CAEDH8BBB7A363.

\subsection{Capacitors}

A digital circuit like an LED driver can change its current consumption significantly within a
short amount of time. Because the connections to the power supply are relatively long, they have 
a substantial parasitic inductivity. So when the IC suddenly needs more current because all its
LEDs have just been switched on, this current is not available instantly. Worse yet, if the LEDs
are on and then are turned off, the current can't immediately stop flowing and will induce
dangerous voltage spikes.

In order to prevent those two situations from happening, I used capacitors. Namely, each LED 
driver has its own \SI{1}{\micro\farad}  ceramic capacitor.

\subsection{Motor}

In order for the whole device to rotate, I needed a motor. Luckily, I could salvage an appropriate
part from an old RC motor boat. It's a big and heavy brushed motor. The fact that it has a built-in
fan and was water-cooled in the boat leads me to think that its power output is much higher than 
what I need.

Because of its high torque, I could attach the motor directly to the rotating shaft instead of 
using gears or a belt drive. That way there are less moving parts and the motor can turn at a lower
rotational speed, both resulting in less wear and noise.

\subsection{Power Transmission}

To transmit power from the stationary power supply to the rotating part, I used slip rings. I made
a relatively simple version, consisting of two round pieces of copper sheet on the rotating end, 
and two long pieces of copper sheet on the stationary part.

The slip rings must be able to transmit the current necessary to light all LEDs up at once.
Assuming that each LED needs 20 mA, this current is about 9.6 A. Additionally, there is a small 
current draw by the microcontroller, but it will be negligible in comparison. Testing has shown 
that the slip rings don't heat up noticeably when conducting 10 A for several minutes.

In practice, however, this current is not reached since the vast majority of pixels are 
switched off when displaying a typical 3D image.

Because two pieces of copper sheet rubbing against each other aren't the most reliable electrical 
connection, I used a \SI{2200}{\micro\farad} electrolytic capacitor to smooth out the supply voltage. Even with it,
I found that the pressure between the two copper conductors had to be quite high for them to form
a good contact.

\newpage
\section{Circuit Diagram}

\begin{figure}[h]
\vspace{4mm}
\includegraphics[width=\textwidth]{./images/schematic.png}
\vspace{-8mm}
\caption{The finished circuit diagram}
\vspace{6mm}
\end{figure}

The circuit diagram --- as well as the PCB layout --- were created in Eagle. Its free version is
limited to two-layer boards of sizes up to \SI{10}{\centi\meter} by \SI{8}{\centi\meter}, but
that's more than enough for my purposes.

In figure 2, the final version of the schematic is pictured. For the sake of simplicity and 
readibility, the layout corresponds roughly to how the board will look.

I had to configure the footprints for the LED drivers and the LEDs by myself, which I did by
reading about the physical dimensions in the respective data sheet. For the resistors and 
capacitors, Eagle had pre-built footprints which I could use.

\subsection{Data Interface}

For transmitting data to the LED drivers, I used a simple row of 5 pins. Two of those pins are the
data lines: The input is connected to the first chip's input, and the output comes from the last
chip's output. The rest of the pins (clock, output enable and latch enable) are control lines,
which go to all three chips simultaneously. The order of these connections was determined when 
making the PCB layout.

\subsection{LED Drivers}

The LED drivers are arranged in a so-called daisy chain configuration, where the serial data output
of each chip is connected to the serial data input of the next chip. This is not only the case 
within one board, but also between the different boards: The data output of one board is connected
to the next board's input. That way, all of the 30 LED drivers can be controlled by a single SPI
bus without anything like a chip select signal. To send data to the chips, the microcontroller 
can simply send out 480 bit without interruption.

\def \iout {I\textsubscript{OUT}}
\def \rext {R\textsubscript{EXT}}

The external resistor which is connected to each LED driver sets the output
current. The following equation, taken from page 15 in the datasheet, describes the relation
between the output current (\iout) and the resistor's value (\rext):

% Lots of ugly hacks to have a non-italic font
\begin{ownmath}
\textrm{\iout} = 15 * \frac{\SI{1.25}{\volt}}{\si{\rext}}
\end{ownmath}

Let's find out the resistance necessary for the output current to be \SI{20}{\milli\ampere}:

\begin{ownmath}
\textrm{\rext} = 15 * \frac{\SI{1.25}{\volt}}{\SI{20}{\milli\ampere}} = \SI{937.5}{\ohm}
\end{ownmath}

The next smaller value within the E24 series is \SI{910}{\ohm}. By choosing this resistance, the 
output current is \SI{20.6}{\milli\ampere}, which is close enough to the desired 
\SI{20}{\milli\ampere}. I ordered 40 through-hole resistors of this value, because at the time I
had in mind to etch my own boards at home.

\subsection{LEDs}

The LEDs have their common anodes connected to \vdd. The cathodes are connected to a driver output
each, and are arranged red - green - blue (from left to right). Series resistors aren't needed 
because the LED drivers handle current regulation.

\section{PCB Design}

\begin{figure}[h]
\vspace{4mm}
\includegraphics[width=\textwidth]{./images/board.png}
\vspace{-4mm}
\caption{The PCB layout in Eagle. Red = upper copper layer, blue = lower copper layer,
green = vias, grey = silkscreen.} 
\vspace{6mm}
\end{figure}

\subsection{General Remarks}

The defining factor for the board layout was the pin layout of the LED drivers. The outputs are all
at the bottom of the chip, while the data interface is at the top. This led me to the layout I have
now: the LEDs are in a row at the bottom of the board where they can be directly connected to the driver 
outputs on the upper copper layer. The control lines are is routed on the back side of the board
behind the chips, with vias going to the front side to connect to each chip. The data lines go 
straight from one chip to the next one on the upper copper layer, with the last output going back
to the board's data interface on the back side. The resistors are placed on the right side of their
respective driver IC, jumping across the data traces.

The supply lines are placed beneath the row of LEDs. Had I placed the supply on the upper edge of the 
board, I still would've had to connect all the 16 anodes to \vdd. The driver chips, on the other hand, 
only need one connection to \vdd and GND each, so by placing the supply lines at the bottom of the board 
I have less traces running across the board vertically.

From a layout perspective, it would've made sense to put the GND line at the top of the board, since the
chips need one connection to that trace each, while the LEDs don't. However, I opted to put both supply
lines right above each other to reduce electromagnetic interference. If all the LEDs are switched on
suddenly and a lot of current starts to flow, a magnetic field builds up quickly around the supply traces.
This could induce bothering or even dangerous currents in the surrounding electronics. By having the two
supply lines close to each other, their magnetic fields largely cancel each other out. Another advantage
of having the supply at the bottom is that the distance to the delicate chips are a bit farther away.

\subsection{Space-saving measures}

In order for the production to be as cheap as possible as well as to minimise the rotating mass, I tried
to make the boards as small as possible. Ultimately, I managed to bring the size down to 
\SI{73}{\milli\meter} by \SI{21}{\milli\meter}

\end{document}