### 2016-05-14 paper.tex
Added & finished chapter 3

\documentclass[a4paper, 11pt, titlepage]{report}

\usepackage[margin=2.0cm]{geometry}

\usepackage{xspace}
\def \iZc {I\textsuperscript{2}C\xspace}
\def \vdd {V\textsubscript{DD}}

% Indent the first line of the first paragraph like the first lines of subsequent paragraphs
\usepackage{indentfirst}

% Don't indent any first lines
\setlength{\parindent}{0mm}

% A little space between paragraphs
\setlength{\parskip}{1.5mm}

% Display
% 1. <chapter title>
% instead of
% Chapter 1
% <chapter title>
\makeatletter
\renewcommand{\@makechapterhead}[1]{%
\vspace*{50 pt}%
{\setlength{\parindent}{0pt} \raggedright \normalfont
\bfseries\Huge\thechapter.\ #1
\par\nobreak\vspace{40 pt}}}
\makeatother

% In order to include images
\usepackage{graphicx}

% Label figures 1, 2, 3, 4... Instead of 1.1, 1.2, 2.1, 3.1 etc.
\usepackage{chngcntr}
\counterwithout{figure}{chapter}

\usepackage{siunitx}

% To have text wrapped around images 
\usepackage{wrapfig}

% Some parameters for wrapfig
\setlength{\intextsep}{0pt}%
\setlength{\columnsep}{15pt}%

% Same as regular math but with a bit of space before and after
\newenvironment{ownmath}
{\vspace{2mm}\hspace{15pt}\begin{math}}
{\end{math}\vspace{2mm}}

\begin{document}

\title{Development of a 3D display}
\author{Balduin Dettling}

% Generate title
\maketitle

% Introduction
% * after chapters to not number them
% TODO: don't show page numbers
\chapter*{Preface}

\pagenumbering{Roman}

\section*{Motivation}

For as long as I can remember, I've been fascinated by all kinds of technology. 
It was clear to me that my final thesis would be a practical one in the field of engineering.
However, I didn't have anything specific in mind for quite some time.

One day, I stumbled upon a project report on the internet about a so-called persistence of 
vision display. It was a simple row of LEDs, driven by a microcontroller, and mounted on an 
electric motor. The contraption was (and presumably still is) able to draw images into the 
air by precisely controlling the row of LEDs. Fascinated by this, I did some research and 
found that a plethora of similar projects exists, some even with detailed instructions on how 
to build the device in question.

I had the idea of building several rows of LEDs and arranging them on top of each other. 
That way, I could display several two-dimensional bitmap images above another, resulting in a
three-dimensional image.

Having had no previsous experience in the field of electronics, it seemed a bit daunting to 
build a 3D display. I could, however, motivate myself to attempt it anyway by telling myself
that the project wouldn't be much more complex than all of the 2D versions out there (which 
turned out to be not very far from the truth).

\section*{Credits}

First and foremost, I would like to express my gratitude to Mr. Patrick Spengler, my physics
teacher and the mentor of this thesis. Not only did he solve numerous malfunctions, but he also
is partly responsible for my interest in electronics and engineering.

I am thankful to Ms. Ursula Schamberger, teacher of the woodworking class, where I built the
wooden frame. Likewise, I would like to thank Mr. Hanspeter Rieder, who kindly granted me
access to the school's electronics workshop.

Last, but most certainly not least, I am indebted to Prof. Heinz Domeisen. He mentored the
advancement and refinement of the thesis in preparation for the national competition organised
by "Swiss Youth in Science".

\section*{Declaration of authorship}

I hereby declare that the following thesis has been authored entirely by myself, Balduin
Dettling, and confirm that all external sources of information are cited correctly.

\vspace{10mm}
\noindent Location and Date: \hspace{0.35\textwidth} Signed:\\
\rule[3mm]{\textwidth}{0.5pt}

% Generate table of contents
\tableofcontents

\chapter{Introduction}

\pagenumbering{arabic}

\section{Current State of The Art}

% todo: references
As discussed before, there is no shortage of projects that are similar to mine, but only
display 2D images. The name "POV display", short for persistence of vision display, has become
commonplace. Another popular name is "Propeller Clock" --- Propeller because it rotates like one,
and clock because the round shape is ideal for displaying an analog clock.

As far as I know, however, only two people have built a working version of a three-dimensional
POV display so far (and have published it on the internet).\footnote{It turned out that there was
a third one: When I posted a picture of my project on the internet, a reddit user told me about
the display he built years before.}
Neither of these is being produced on a grand scale, instead they are --- like most 2D versions
--- side projects of makers or engineers.

\section{Goals}

The objective of this project is to build a working prototype of a 3D persistence of vision
display and to document the process of doing so. The necessary steps are (roughly) as follows:

\begin{itemize}
	\item Understand and explain how and why such a display can work in principle
	\item Come up with realistic technical specifications, founded on the previously outlined 
theory  
	\item Select parts and build the device according to the specifications
	\item Write a program that makes the device display a 3D image
	\item Draw conclusions: Did I achieve my goals? What could have been done better? 
How can the project be further improved?
\end{itemize}

\chapter{Theory and Planning}

\section{Principle of Operation}

\subsection{Persistence of Vision}

Persistence of Vision refers to the optical illusion that results whenever the image we look
at quickly changes. In such situations, we don't immediately see the new image, instead we
continue seeing the old image for a short timespan after it has vanished. If wikipedia is to 
be believed, this timespan is about $\frac{1}{25}$ of a second, but of course it is going to 
be different for each pair of eyes.

Although we don't normally see this effect in action, there are situations where it becomes
apparent. For example, if a car drives by at nighttime, we see traces of light behind the
actual car lights for a short moment (Assuming our eyes don't follow the car and stay still
relative to the environment). In general, bright flashes of light followed by (relative)
darkness tend to be the most pronounced manifestation of persistence of vision. It is exactly 
this situation that can be recreated and exploited for the successful operation of a persistence
of vision display.

\subsection{Exploiting Persistence of Vision}

Normally, Persistence of Vision is nothing more than an interesting effect. However, we can 
use this weakness of the human eye to our benefit.

If a light source is spun quickly enough in a circle, we don't see a spinning
point anymore, but instead we see one solid circle. This is because the rotational period is
shorter than the time it takes for the light to fade away on our retinas. The rotational period
at which we stop seeing a flickering circle and start seeing a solid, still circle is called
the flicker fusion threshold.

By controlling this light source quickly and accurately, we can make it so it is always turned 
on at certain points in the circle, and always turned off at others. By doing so, we are
effectively multiplexing a zero-dimensional display (a point-like light source, e.g. an LED)
so that it displays a one-dimensional image. By rotating a whole row of LEDs (a one-dimensional
display), we can display two-dimensional images. Finally, if there are multiple rows of LEDs
that together form a two-dimensional display, it is possible to display a three-dimensional
image.

\section{Specifications}

\begin{itemize}

	\item Resolution of 16 * 100 * 10 pixels (radius * circumference * height)
	\item Each pixel consists of an RGB LED
	\item Colour depth of 3 bit per pixel, or one bit per LED
	\item Rotational frequency of at least 30 Hz, or as fast as it needs to be in order to not flicker

\end{itemize}

\section{Components}

\subsection{Control}

For controlling the LEDs, I have bought a Teensy 3.1 microcontroller development board. It has a
small footprint --- ideal for the fast rotation --- and still packs quite some power: The processor
runs at 96 MHz at a register width of 32 bit. There are 34 configurable I/O pins.

The controller has to have some kind of reference point, so that it can calculate the current speed
and angular position of the device. For this, I used a simple hall sensor on the rotating side, 
which passes by a stationary magnet after each rotation.

\subsection{Data Transmission}

34 I/O pins are quite a lot, but not quite enough to directly control 480 LEDs. In order to still
access all those LEDs individually, I had to use LED drivers. These come in the form of integrated
circuits and receive some amount of data that they use to control a corresponding amount of LEDs.

Initially, I had planned to implement a colour depth of 8 bits, allowing for 
$2^{24} = 16777216$ different colours. However, there were several difficulties that I would have
had to deal with. First of all, all of the available LED drivers had one or both of the following
shortcomings:

\begin{itemize}

	\item Controlled using \iZc  instead of SPI

Because \iZc uses pullup resistors instead of a push-pull drive, it takes a little moment for the
logic level to return to its normal state. In order to ensure correct data transmission, it is 
necessary to set a relatively slow clock speed. The two \iZc interfaces of the Teensy only support
clock speeds of 2.4 MHz --- and because \iZc requires some overhead data to send out the slave's
addresses, the actual data rate will be even lower.

However, the minimum data rate required to control the LEDs at an 8-bit colour depth is as follows:

\begin{ownmath}
% Yes it's a mess but \si makes it a unit, so it is not in cursive while still being small enough
24\,\frac{\si{bit}}{\si{pixel}} * 16000\,\frac{\si{pixels}}{\si{rotation}} * 30\,\frac{\si{rotations}}{\si{second}} = \SI{11.52}{Mbps} 
\end{ownmath}

This means that I'd need at least 5 parallel \iZc lines to control all the LEDs with 8 bit colour
depth. Because each Teensy only has two of them, I'd need three Teensy boards, which would be
possible but very cumbersome.

	\item Insufficient PWM frequency

If the colour depth is more than one bit, the LEDs will have to be dimmed using PWM. Normally, a
rather low frequency of a few hundred to a few thousand Hz is sufficient as a PWM frequency, and
most LED drivers modulate at frequencies within that range. In my case, however, the LEDs are 
moving into a new pixel 3000 times a second (assuming rotationat 30 Hz). This means that a PWM
frequency under 3000 Hz will involve losing information, and one slightly above 3000 Hz still
won't look good. There would have to be a significant number of PWM cycles within each pixel,
for example 10, which would require a PWM frequency of 30 KHz.

\end{itemize}

Because of that, I ditched the idea of 8-bit colour depth and instead went for the simpler approach
of one-bit colour depth. Both problems are solved by switching to 1 bit: The data rate doesn't have
to be as high (and there are lots of chips that are controlled by SPI), and PWM is not necessary
anymore.

% todo: Find out if this is necessary in the end
\newpage
\subsection{LED drivers}

\begin{wrapfigure}{r}{0cm}

\includegraphics[scale=0.35]{./images/tlc-pins.png}
\caption{Pin layout of the TLC5927 LED driver}

\end{wrapfigure}

After quite some searching around, I decided to use the TLC5927 by Texas Instruments. This is
essentially a shift register with some extra features. Like every shift register, it has a serial
data input (SDI), a serial data output (SDO), and a clock pin (CLK). Additionally, there is a latch
pin (LE), which serves the purpose of only refreshing the state of the LEDs once all data has been
sent.

Furthermore, there is a pin called "output enable" (OE). If it is at GND, all LEDs are turned off,
and it's at \vdd, they are controlled by the data that the driver chip has received.

The most important feature, however, is the fact that each output is controlled by a constant
current driver. Without these, it would've been necessary to solder a resistor in series
with each individual LED, for a total of 480 resistors, 960 additional solder joints and a whole
lot of wasted board space. With the constant current drivers, however, I could use a single
resistor to set the current for all 16 LEDs attached to one chip.

Unlike most simple shift registers, the drivers' outputs are current sinks instead of current 
sources. This will be important when selecting the LEDs: In order to control each colour separately 
in an RGB LED, the cathodes have to be separated, while the anodes can be connected together.

\subsection{LEDs}

Like all the other parts, I ordered the LEDs from Digikey. Their website
contains an excellent tool, enabling the user to restrict the selection based on different criteria.
My requirements were the following:

\begin{itemize}

	\item Colours: Red, green and blue
	\item Current: 20 mA or more. Because the LEDs are within one pixel only $\frac{1}{100}$ of the
time, they are essentially pulse width modulated, reducing their effective brightness. In order to
still see the image clearly without darkening the room, they have to be bright enough.
	\item Mounting type: Surface mount, so that I can route traces past the LEDs on the other side 
of the PCB
	\item Size: At most 5 * 5 mm, preferably smaller, but still large enough to be able to solder them
by hand
	\item Diffused lens, so the colours mix well even when viewing the LED directly instead of using
it for illumination.

\end{itemize}

These requirements reduced the category "LED Indication - Discrete" from 18940 to 130 products.
From those remaining products, I chose the one that was cheapest in a quantity of 160. This 
led me to an LED with the beautiful name CLVBA-FKA-CAEDH8BBB7A363.

\subsection{Capacitors}

A digital circuit like an LED driver can change its current consumption significantly within a
short amount of time. Because the connections to the power supply are relatively long, they have 
a substantial parasitic inductivity. So when the IC suddenly needs more current because all its
LEDs have just been switched on, this current is not available instantly. Worse yet, if the LEDs
are on and then are turned off, the current can't immediately stop flowing and will induce
dangerous voltage spikes.

In order to prevent those two situations from happening, I used capacitors. Namely, each LED 
driver has its own \SI{1}{\micro\farad}  ceramic capacitor.

\subsection{Motor}

In order for the whole device to rotate, I needed a motor. Luckily, I could salvage an appropriate
part from an old RC motor boat. It's a big and heavy brushed motor. The fact that it has a built-in
fan and was water-cooled in the boat leads me to think that its power output is much higher than 
what I need.

Because of its high torque, I could attach the motor directly to the rotating shaft instead of 
using gears or a belt drive. That way there are less moving parts and the motor can turn at a lower
rotational speed, both resulting in less wear and noise.

\subsection{Power Transmission}

To transmit power from the stationary power supply to the rotating part, I used slip rings. I made
a relatively simple version, consisting of two round pieces of copper sheet on the rotating end, 
and two long pieces of copper sheet on the stationary part.

The slip rings must be able to transmit the current necessary to light all LEDs up at once.
Assuming that each LED needs 20 mA, this current is about 9.6 A. Additionally, there is a small 
current draw by the microcontroller, but it will be negligible in comparison. Testing has shown 
that the slip rings don't heat up noticeably when conducting 10 A for several minutes.

In practice, however, this current is not reached since the vast majority of pixels are 
switched off when displaying a typical 3D image.

Because two pieces of copper sheet rubbing against each other aren't the most reliable electrical 
connection, I used a \SI{2200}{\micro\farad} electrolytic capacitor to smooth out the supply 
voltage. Even with it, I found that the pressure between the two copper conductors had to be quite
high for them to form a good contact.

\newpage
\section{Circuit Diagram}

\begin{figure}[h]
\vspace{4mm}
\includegraphics[width=\textwidth]{./images/schematic.png}
\vspace{-8mm}
\caption{The finished circuit diagram}
\vspace{6mm}
\end{figure}

The circuit diagram --- as well as the PCB layout --- were created in Eagle. Its free version is
limited to two-layer boards of sizes up to \SI{10}{\centi\meter} by \SI{8}{\centi\meter}, but
that's more than enough for my purposes.

In figure 2, the final version of the schematic is pictured. For the sake of simplicity and 
readibility, the layout corresponds roughly to how the board will look.

I had to configure the footprints for the LED drivers and the LEDs by myself, which I did by
reading about the physical dimensions in the respective data sheet. For the resistors and 
capacitors, Eagle had pre-built footprints which I could use.

\subsection{Data Interface}

For transmitting data to the LED drivers, I used a simple row of 5 pins. Two of those pins are the
data lines: The input is connected to the first chip's input, and the output comes from the last
chip's output. The rest of the pins (clock, output enable and latch enable) are control lines,
which go to all three chips simultaneously. The order of these connections was determined when 
making the PCB layout.

\subsection{LED Drivers}

The LED drivers are arranged in a so-called daisy chain configuration, where the serial data output
of each chip is connected to the serial data input of the next chip. This is not only the case 
within one board, but also between the different boards: The data output of one board is connected
to the next board's input. That way, all of the 30 LED drivers can be controlled by a single SPI
bus without anything like a chip select signal. To send data to the chips, the microcontroller 
can simply send out 480 bit without interruption.

\def \iout {I\textsubscript{OUT}}
\def \rext {R\textsubscript{EXT}}

The external resistor which is connected to each LED driver sets the output
current. The following equation, taken from page 15 in the datasheet, describes the relation
between the output current (\iout) and the resistor's value (\rext):

% Lots of ugly hacks to have a non-italic font
\begin{ownmath}
\textrm{\iout} = 15 * \frac{\SI{1.25}{\volt}}{\si{\rext}}
\end{ownmath}

Let's find out the resistance necessary for the output current to be \SI{20}{\milli\ampere}:

\begin{ownmath}
\textrm{\rext} = 15 * \frac{\SI{1.25}{\volt}}{\SI{20}{\milli\ampere}} = \SI{937.5}{\ohm}
\end{ownmath}

The next smaller value within the E24 series is \SI{910}{\ohm}. By choosing this resistance, the 
output current is \SI{20.6}{\milli\ampere}, which is close enough to the desired 
\SI{20}{\milli\ampere}. I ordered 40 through-hole resistors of this value, because at the time I
had in mind to etch my own boards at home.

\subsection{LEDs}

The LEDs have their common anodes connected to \vdd. The cathodes are connected to a driver output
each, and are arranged red - green - blue (from left to right). Series resistors aren't needed 
because the LED drivers handle current regulation.

\section{PCB Design}

\begin{figure}[h]
\vspace{4mm}
\includegraphics[width=\textwidth]{./images/board-layout.png}
\vspace{-4mm}
\caption{The PCB layout in Eagle. Red = upper copper layer, blue = lower copper layer,
green = vias, grey = silkscreen.} 
\vspace{6mm}
\end{figure}

\subsection{General Remarks}

The defining factor for the board layout was the pin layout of the LED drivers. The outputs are all
at the bottom of the chip, while the data interface is at the top. This led me to the layout I have
now: the LEDs are in a row at the bottom of the board where they can be directly connected to the 
driver outputs on the upper copper layer. The control lines are is routed on the back side of the 
board behind the chips, with vias going to the front side to connect to each chip. The data lines 
go straight from one chip to the next one on the upper copper layer, with the last output going 
back to the board's data interface on the back side. The resistors are placed on the right side of 
their respective driver IC, jumping across the data traces.

The supply lines are placed beneath the row of LEDs. Had I placed the supply on the upper edge of 
the board, I still would've had to connect all the 16 anodes to \vdd. The driver chips, on the 
other hand, only need one connection to \vdd and GND each, so by placing the supply lines at the 
bottom of the board I have less traces running across the board vertically.

From a layout perspective, it would've made sense to put the GND line at the top of the board, 
since the chips need one connection to that trace each, while the LEDs don't. However, I opted to 
put both supply lines right above each other to reduce electromagnetic interference. If all the 
LEDs are switched on suddenly and a lot of current starts to flow, a magnetic field builds up 
quickly around the supply traces. This could induce bothering or even dangerous currents in the 
surrounding electronics. By having the two supply lines close to each other, their magnetic fields 
largely cancel each other out. Another advantage of having the supply at the bottom is that the 
distance to the delicate chips are a bit farther away.

\subsection{Space-saving measures}

In order for the production to be as cheap as possible as well as to minimise the rotating mass, I 
tried to make the boards as small as possible. Ultimately, I managed to bring the size down to 
\SI{73}{\milli\meter} by \SI{21}{\milli\meter}.

In order to achieve this, but also to avoid large gaps between the pixels, I placed the LEDs as 
close as possible to each other. The datasheet claims a width of \SI{2.8 +- 0.2}{\milli\meter}, 
so I decided to place the LEDs at \SI{3}{\milli\metre} increments. On a grid of \SI{12.5}{mil}, 
the next larger distance is \SI{125}{mil}, which is equal to \SI{3.175}{mm}.

The connections to the cathodes of the red LED needs to pass by the other two cathodes, since 
they're located on the lower side of the LED package. Because the LEDs are so close to each other, 
the only option was to use the lower copper layer. The only place to go back to the upper copper 
layer was right beneath the driver chips.

Initially, I had placed the capacitors right above the ICs. However, I could save a small bit of
vertical board size by placing them at an offset to the right. That way I could push them down 
until where the pins begin, which is a bit further down than the edge of the chip itself.

\subsection{Thermal Considerations}

A positive side effect of placing some vias beneath the chip is their thermal conductivity. That
way, the chips' waste heat can be dissipated more effectively. The maximum power output of the 
chips --- when all the attached LEDs are turned on --- is as follows:

\begin{ownmath}
\textrm{P\textsubscript{MAX}} = (\SI{5}{\volt} - \textrm{V\textsubscript{F}}) * 
\SI{20.6}{\milli\ampere\per output} * \SI{16}{outputs} = \SI{867.9}{\milli\watt}
\end{ownmath}

V\textsubscript{F} is different depending on the colour of the LED, so I just used the average
value, which is \SI{2.633}{\volt}. The actual values are \SI{1.9}{\volt} for red and \SI{3}{\volt}
for green and blue.

A power output of almost a watt is not to be underestimated, and tests have shown that that the
boards do get quite hot when operated at full power. In reality, however, the situation is much
less severe because the LEDs are typically turned on only during a fraction of the rotation.
Furthermore, the quick rotation results in excellent air circulation, which aids the dissipation
of heat.

\chapter{Fabrication}

\section{Circuit Boards}

\subsection{Fabrication}

There's no shortage of firms who produce PCBs according to a custom layout. One of the cheapest 
providers to ship small production runs within a reasonable timeframe and with good production
quality turned out to be OSHPark. They only produce in multiples of 3, so I ordered 12 instead of
10 (which I'd have done anyway to have some spare boards). I submitted my order on the 5th of 
October 2014 and paid US\$ 47. Ten days later I received a message that the boards were finished,
and after another ten days of waiting they arrived at my doorstep.

\begin{figure}[h]
\vspace{3mm}
\includegraphics[width=\textwidth]{./images/board-finished.jpeg}
\vspace{-8mm}
\caption{A finished circuit board} 
\end{figure}

\subsection{Assembly}

As soon as I got the boards in the mail, I started soldering the parts to them. I started with the
LEDs to avoid having to solder in the tight gap that would later be formed between the LEDs and
their driver chips. It took me two hours to solder the first board, but with increasing practice
the time for one row of LEDs got down to 25 or 30 minutes.

The chips were a lot easier to solder than the LEDs, despite the distance between the individual
pins only being \SI{0.635}{\milli\meter}. I could simply drown all the pins in solder, and then 
wick away all the solder bridges. That way I could solder twelve pins within the blink of an eye.

\subsection {Testing and Debugging}

With the boards finished, I wanted to know whether or not they work. In order to find out, I wrote
a little program on the microcontroller to light the LEDs up in all possible colours consecutively.
It did not work at the first try, so I used an oscilloscope to see if data was being sent at all. 
This was the case, the SPI library on the Teensy performed its task flawlessly.

It turned out that the culprit were the bad contacts. Instead of soldering the wires onto the board
and the microcontroller, I simply inserted them into the respective holes. This caused the 
connection to have too much parasitic capacitance, and the signals (which were being sent at 4 MHz)
did not have time to propagate up to their destination.

I soldered the cables on properly and sure enough, the board worked as intended. I repeated this
procedure for all ten boards to see if they all worked as well as the first one, which wasn't the 
case.

Because of reasons unbeknownst to me, one particular LED driver started turning on all of its LEDs
at a moderate brightness as soon as I powered it up --- even though no data was being sent at all.
At first I thought that the serial data input had a solder bridge pulling it up to \vdd, and
immediately sending a whole lot of 1's upon powering up the board. However, the data input pin was
not close enough to a positive supply line for this to be an issue. Additionally, there would have 
to be a clock signal as well as a pulse on the latch pin for the LEDs to light up. I gave up and 
assumed that the chip was faulty, and when I replaced it with a spare part everything worked 
perfectly.

Apart from that, there were no major issues. I had a few LEDs which didn't light up when they 
should have. One of them had a solder bridge between the common anode and the cathode of the red
LED, which made a voltage drop across the red LED impossible. Two were simply soldered on poorly
and weren't connected to the drivers at all. A few other LEDs didn't work and I found no obvious
issues. I replaced them with spares, which did the trick.

\section{Physical Assembly}

\subsection{Frame}

The frame was built out of beech wood strips with a square cross section and a side length of 
\SI{1.5}{\centi\meter}. The baseplate is made out of plywood and has the dimensions 
\SI{15}{\centi\meter} by \SI{15}{\centi\meter} by \SI{1}{\centi\meter}. On it, I mounted two 
vertical wood strips with a length of \SI{20}{\centi\meter}, which were both stabilise by three
struts. A horizontal board, mounted to the vertical strips at a height of \SI{20}{\centi\meter}, 
holds the motor.

On the top of the two vertical strips I placed another strip to hold the shaft in its position. 
A simple hole served as a bearing, which I later replaced with a plastic insert (kindly sponsored
by IGUS).

\subsection{Slip Rings}

In order to supply the rotating parts with electricity, I mounted a wooden disc right at the bottom
of the shaft. On both sides of that disc I glued on a round piece of copper sheet, which I then
connected to cables going up to where the electronics would later be.

On the stationary end, two long pieces of copper sheet serve as contacts. Because of their 
spring-like properties, they exert a certain amount of pressure on the rotating disc.

Despite this, a considerable amount of pressure was necessary to form a good contact. The contact
pressure proved to be a primitive, but usable way of roughly adjusting the rotational speed of
the display. Sadly it's quite an inefficient approach: The motor uses a lot of current when 
friction is applied to its load. It becomes noticeably warm after a few minutes of use, and 
measurements on the current limiting resistors have shown that it draws about \SI{6.5}{\ampere} at
\SI{5}{\volt}.

This constant pressure, in addition to the fact that copper is quite a soft metal, results in 
considerable wear on the slip rings. Luckily, the stationary ends can be replaced without much 
hassle. The rotating part would be a pain to replace, but since wear is distributed across the 
whole copper disc, it will be a long time before this needs to be done.

\subsection{Controller}

The microcontroller is mounted on another wooden disc, located a few centimetres above the slip
rings. It also hosts the hall sensor, which is used to measure the rotational speed and calculate
the position of the rotating part at any given moment. The corresponding magnet is mounted on a 
piece of threaded rod, which itself is mounted on the same board as the motor.

Additionally, the \SI{2200}{\micro\farad} capacitor is located on the upper wooden disc. In the 
end, this place has become a bit of a mess, especially with all the cables going up to the circuit 
boards. Luckily this isn't visible while the apparatus is running.

\subsection{Circuit Boards}

The boards were already equipped with components, but what was missing was the hold to mount them
on the shaft. The diameter of this shaft is \SI{4}{\milli\meter}. In order to find a good hole 
size, I drilled a few test holes on a spare board. A hole of \SI{4}{\milli\meter} was so small for
the boards to fit on the shaft, while \SI{4.5}{\milli\meter} was obviously too big. After 
chamfering the top end of the shaft, the board with a \SI{4}{\milli\meter} hole just barely could
slide on the shaft. This tight fit proved to be advantageous, because torque can now be transmitted
reliably from the shaft to the boards, and they don't slide around.

\chapter{Programming}

\end{document}